---
title: "Home range estimation"
format:
  html:
    code-fold: true
    fig-dpi: 300
---

```{r r-setup, warning = FALSE, message = FALSE}
#| code-summary: "R setup"

# packages
library(tidyverse)
library(kableExtra)
library(ggpubr)

# options
options(
  dplyr.summarise.inform = FALSE,
  knitr.table.html.attr = "quarto-disable-processing=true"
)

# define site and population colors
site_colors <- c("HE" = "royalblue", "LE" = "darkorange")
population_colors <- c(
  "HE 2023" = "royalblue",
  "LE 2023" = "darkorange",
  "LE 2024" = "goldenrod4"
)
```

## Mark-recapture dataset structure

Here we load the compiled longitudinal mark-recapture dataset (`mr_data.rds`) and inspect its structure, including individual identifiers, capture timestamps (year and day of the year), morphometric attributes (snout-vent length, mass), sex, and spatial coordinates.

```{r load-and-show-mr-data}
#| code-fold: false
mr_data <- readRDS("data/mr_data.rds")
head(mr_data)
str(mr_data)
```

## Probability of recapturing adult lizards

To assess capture saturation and sampling completeness across field seasons, we filter the dataset for adult lizards (mean snout-vent length $> 75\text{ mm}$) and compute the cumulative number of unique individuals captured as a function of the day of the year (DOY) across sites and years.

```{r plot-cummulative-captures}
#| code-summary: "Plot cummulative captures"
#| fig-width: 7
#| fig-height: 4
#| fig-align: center
#| out-width: 80%

# filter adults
adults_by_site_year <- mr_data |>
  group_by(site, year, sex, id) |>
  summarise(avg_svl = mean(svl, na.rm = TRUE)) |>
  filter(avg_svl > 75) |>
  ungroup()

# filter mark-recapture data to include only adults
mr_data_adults <- mr_data |>
  semi_join(adults_by_site_year, by = c("site", "year", "sex", "id"))

# plot
mr_data_adults |>
  filter(
    year > 2022,
    paste(site, year) != "HE 2025"
  ) |>
  group_by(site, year, sex, id) |>
  mutate(capture_number = row_number()) |>
  ungroup() |>
  group_by(site, year, doy) |>
  summarise(
    n_first_capture = sum(capture_number == 1), .groups = "drop"
  ) |>
  mutate(
    cummulative_captures = cumsum(n_first_capture)
  ) |>
  ggplot(aes(x = doy, y = cummulative_captures, color = site)) +
  geom_line(linewidth = 1) +
  facet_wrap(~year, scales = "free_x") +
  labs(
    x = "Date",
    y = "Number of unique adult lizards"
  ) +
  scale_color_manual(values = site_colors) +
  scale_x_continuous(
    breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335),
    labels = month.abb,
  ) +
  theme_minimal() +
  theme(
    panel.border = element_rect(fill = NA, linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.25),
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    legend.position = "top",
    strip.text = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linewidth = 0.25, color = "gray80")
  )

```

Saturation is achieved across sites and years, indicating that the majority of adult lizards have likely been captured at least once during the field seasons.

## Inter-annual survival probabilities

To evaluate population persistence across seasons, we compute empirical apparent survival probabilities for adult males by tracking whether an individual detected in a given year was recaptured in the subsequent year. The resulting annual survival rates are plotted across consecutive yearly transitions for high- and low-elevation populations.

```{r plot-inter-annual-survival-probabilities}
#| code-summary: "Plot inter-annual survival probabilities"
#| fig-width: 4
#| fig-height: 4
#| fig-asp: 1
#| fig-align: center
#| out-width: 50%

mr_data_adults |>
  select(site, id, sex, year) |>
  distinct() |>
  group_by(site, sex, id) |>
  mutate(all_years = list(year)) |>
  rowwise() |>
  mutate(
    surv_next_year = as.numeric((year + 1) %in% all_years)
  ) |>
  ungroup() |>
  select(-all_years) |>
  group_by(site, sex, year) |>
  summarise(
    surv_next_year = mean(surv_next_year), .groups = "drop"
  ) |>
  filter(year < 2025) |>
  filter(year > 2022) |>
  filter(sex == "M") |>
  mutate(year = as.integer(year)) |>
  mutate(surv_next_year = as.numeric(surv_next_year)) |>
  ggplot(
    aes(
      x = paste(year, "to", year + 1),
      y = surv_next_year,
      color = site
    )
  ) +
  geom_line(aes(group = site), linewidth = 1) +
  geom_point(size = 3) +
  labs(
    x = "Interval",
    y = "Mean inter-annual survival"
  ) +
  scale_color_manual(values = site_colors) +
  theme_minimal() +
  theme(
    panel.border = element_rect(fill = NA, linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.25),
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    legend.position = "top",
    strip.text = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(
      linewidth = 0.25,
      color = "gray80"
    )
  )
```

Below we table the inter-annual survival probabilities for adult males across sites and years.

```{r table-inter-annual-survival-probabilities}
mr_data_adults |>
  select(site, id, sex, year) |>
  distinct() |>
  group_by(site, sex, id) |>
  mutate(all_years = list(year)) |>
  rowwise() |>
  mutate(
    surv_next_year = as.numeric((year + 1) %in% all_years)
  ) |>
  ungroup() |>
  select(-all_years) |>
  group_by(site, sex, year) |>
  summarise(
    surv_next_year = mean(surv_next_year), .groups = "drop"
  ) |>
  mutate(surv_next_year = round(surv_next_year, 2)) |>
  filter(year < 2025) |>
  filter(year > 2022) |>
  filter(sex == "M") |>
  select(-sex) |>
  mutate(year = as.integer(year)) |>
  mutate(surv_next_year = as.numeric(surv_next_year)) |>
  rename(
    Site = site,
    Year = year,
    `Inter-annual survival probability` = surv_next_year
  ) |>
  knitr::kable() |>
  kableExtra::kable_styling(
    full_width = FALSE,
    font_size = 16
  )
```

::: {.callout-note}
Inter-annual survival probabilities from 2024 to 2025 are extremely low at high elevation because we did not exhaustively sample that population during 2025 due to logistical constraints.
:::

## Home range estimation

We delineate individual home ranges by computing minimum convex polygons (MCP) via convex hulls (`chull`) using projected UTM coordinates. We retain individuals with at least 4 spatially-reference captures and filter out coordinate points $> 150\text{ m}$ from each individual's mean centroid to remove GPS artifacts and/or potential misidentifications. The resulting polygon coordinates are saved to `data/home_ranges.rds` and visualized with site-mean-centered coordinates to compare spatial extents across elevation sites.

```{r estimate-home-ranges}
#| code-fold: false

home_ranges <- mr_data |>
  filter(if_all(c(lat, lon, utm_x, utm_y), ~ !is.na(.x))) |>
  group_by(site, id, sex) |>
  mutate(
    n_obs = n(),
    mean_utm_x = mean(utm_x),
    mean_utm_y = mean(utm_y),
    dist_to_center = sqrt((utm_x - mean_utm_x)^2 + (utm_y - mean_utm_y)^2)
  ) |>
  filter(
    dist_to_center <= 150,
    n_obs >= 4
  ) |>
  slice(chull(utm_x, utm_y)) |>
  ungroup() |>
  select(site, id, sex, utm_x, utm_y)

saveRDS(home_ranges, file = "data/home_ranges.rds")
```

```{r plot-lizard-home-ranges}
#| code-summary: "Plot lizard home ranges"
#| fig-width: 7
#| fig-height: 4
#| fig-align: center
#| out-width: 80%

home_ranges |>
  group_by(site) |>
  mutate(
    x_dim = utm_x - mean(utm_x),
    y_dim = utm_y - mean(utm_y)
  ) |>
  ungroup() |>
  filter(sex == "M") |>
  ggplot(aes(x = x_dim, y = y_dim, color = site)) +
  geom_polygon(aes(group = id), linewidth = 0.5, fill = NA) +
  labs(
    x = "Distance (m, relative to site mean)",
    y = "Distance (m, relative to site mean)"
  ) +
  theme_minimal() +
  scale_color_manual(values = site_colors) +
  facet_wrap(
    ~site, 
    scales = "free",
    space = "free_x",
    labeller = as_labeller(c(
      HE = "High Elevation (HE)",
      LE = "Low Elevation (LE)"
    ))
  ) +
  scale_x_continuous(
    breaks = seq(-400, 400, by = 100)
  ) +
  theme(
    panel.border = element_rect(fill = NA, linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.25),
    legend.title = element_blank(),
    legend.text = element_text(size = 14),
    strip.text = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
      linewidth = 0.25,
      color = "gray80"
    ),
    legend.position = "none"
  )

```

::: {.callout-note}
The `LE` population is divided between two neighbouring outcrops while the `HE` population extends across a continuous area.
:::