---
title: "Resting metabolic rates"
format:
  html:
    code-fold: true
    fig-dpi: 300
---

```{r r-setup, warning = FALSE, message = FALSE}
#| code-summary: "R setup"

# packages
library(tidyverse)
library(kableExtra)
library(ggpubr)
library(viridis)
library(rstan)
library(bayesplot)
library(tidybayes)

# options
options(
  dplyr.summarise.inform = FALSE,
  knitr.table.html.attr = "quarto-disable-processing=true"
)

# define site and population colors
site_colors <- c("HE" = "royalblue", "LE" = "darkorange")
population_colors <- c(
  "HE 2023" = "royalblue",
  "LE 2023" = "darkorange",
  "LE 2024" = "goldenrod4",
  "TLE 2024" = "maroon"
)
```

## Respirometry dataset structure

```{r respirometry-dataset-structure}
respirometry <- readRDS("data/respirometry.rds")
head(respirometry)
str(respirometry)
```

```{r example-respirometry-data}
#| code-summary: "Example respirometry data for L089 in 2024"
#| fig-align: "center"
#| fig-width: 5
#| fig-height: 4
#| out-width: "66%"

respirometry |>
  mutate(site = paste(site, "E", sep = "")) |>
  filter(year == 2024, id == "L089") |>
  filter(trial_time < 15 * 60) |>
  group_by(temp) |>
  mutate(o2_scaled = o2 - max(o2)) |>
  ggplot(aes(x = trial_time, y = o2_scaled)) +
  geom_point(aes(color = temp, group = temp), alpha = 0.25) +
  geom_smooth(method = "lm", aes(color = temp, group = temp)) +
  scale_color_viridis(breaks = seq(10, 40, by = 10)) +
  labs(
    y = expression(Delta * O[2] ~ "(%)"),
    x = "Trial Time (s)",
    color = "Temp. (°C)"
  ) +
  theme_classic()
```

## Estimating Resting Metabolic Rates

Below we transform the raw respirometry data into resting metabolic rates (RMR) and mass-specific RMR for each individual. This process involves fitting linear models to the oxygen consumption data over time, extracting the slope as the RMR, and then normalizing by body mass to obtain mass-specific RMR. Further, we control-correct for background fluctuations in oxygen consumption by subtracting the slope of the control measurements.

```{r estimating-rmrs}
rmr_data <- respirometry |>
  group_by(site, year, id, svl, mass, temp) |>
  summarise(
    rmr = coefficients(lm(o2 ~ trial_time))[2] * 60, # ml min-1
    rmr_control = coefficients(lm(o2_control ~ trial_time))[2] * 60 # ml min-1
  ) |>
  ungroup() |>
  mutate(
    rmr_mass_specific = - (rmr - rmr_control) / mass, # ml O2 min-1 g-1
  ) |>
  mutate(
    rmr_mass_specific = ifelse(
      rmr_mass_specific < 0, NA, # further removal of anomalus measurements
      rmr_mass_specific
    )
  ) |>
  mutate(log_rmr_mass_specific = log(rmr_mass_specific)) |>
  select(
    site, year, id, svl, mass, temp, rmr_mass_specific, log_rmr_mass_specific
  )
rmr_data
```

The raw RMR data looks as:

:::{.panel-tabset}

## Raw rates

```{r plot-raw-rmr-data}
#| code-summary: "Raw mass-specific RMR estimates"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 5
#| out-width: "66%"

rmr_data |>
  mutate(site = paste(site, "E", sep = "")) |>
  filter(paste(site, year) != "HE 2024") |>
  ggplot(aes(x = temp, y = rmr_mass_specific, color = paste(site, year))) +
  geom_line(aes(group = id), linewidth = 1, alpha = 0.3) +
  facet_wrap(
    ~ paste(site, year),
    axes = "all",
    labeller = as_labeller(function(label) {
      sprintf(
        "<span style='color: %s;'>%s</span>",
        population_colors[label],
        label
      )
    })
  ) +
  scale_color_manual(
    values = population_colors
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    panel.border = element_rect(),
    strip.text = ggtext::element_markdown(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "none"
  ) +
  labs(
    x = "Body temperature (°C)",
    y = expression(paste("Mass-specific RMR (ml O"[2], " g"^{-1}, " min"^{-1}, ")")), # nolint
  )
```

## Log-transformed rates

```{r plot-log-transformed-rmr-data}
#| code-summary: "Log-transformed mass-specific RMR estimates"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 5
#| out-width: "66%"

rmr_data |>
  mutate(site = paste(site, "E", sep = "")) |>
  filter(paste(site, year) != "HE 2024") |>
  ggplot(aes(x = temp, y = log_rmr_mass_specific, color = paste(site, year))) +
  geom_line(aes(group = id), linewidth = 1, alpha = 0.3) +
  facet_wrap(
    ~ paste(site, year),
    axes = "all",
    labeller = as_labeller(function(label) {
      sprintf(
        "<span style='color: %s;'>%s</span>",
        population_colors[label],
        label
      )
    })
  ) +
  scale_color_manual(
    values = population_colors
  ) +
  theme_classic() +
  theme(
    axis.line = element_blank(),
    panel.border = element_rect(),
    strip.text = ggtext::element_markdown(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "none"
  ) +
  labs(
    x = "Body temperature (°C)",
    y = expression(paste("Mass-specific RMR (ml O"[2], " g"^{-1}, " min"^{-1}, ")")), # nolint
  )
```

:::

::: {.callout-note}
The `TLE` population corresponds to individuals from outcrops adjacent to the main `LE` site collected during 2024 to explore seasonal variability in RMR
:::

## Missing RMR trials

```{r plot-missing-rmr-trials}
#| code-summary: "Visualize missing RMR trials"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 7
#| out-width: "66%"

rmr_data |>
  mutate(site = paste(site, "E", sep = "")) |>
  filter(paste(site, year) != "HE 2024") |>
  ggplot(aes(x = temp, y = id, fill = paste(site, year))) +
  geom_tile(color = "lightgray") +
  facet_wrap(~paste(site, year), scales = "free", ncol = 4) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_manual(
    values = population_colors
  ) +
  theme_classic() +
  theme(
    axis.text.y = element_text(size = 10),
    legend.position = "none",
    strip.text = element_text(size = 12, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA),
    axis.ticks = element_line(color = "black"),
    axis.line = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(
    x = "Body temperature (°C)",
    y = "Individual ID"
  )

```

## Estimating intercepts and slopes of RMR

### Model description

#### Specification

We modeled the relationship between log mass-specific metabolic rate ($R$) and body temperature ($T$) using a hierarchical Bayesian framework to account for the nested structure of our data. Specifically, we modeled $R$ as a linear function of $T$ with individual-specific intercepts and slopes:

$$R_{i,j} \sim \mathcal{N}(\alpha_j + \beta_j \cdot T_{i,j}, \sigma)$$

Where $R_{i,j}$ is the log mass-specific metabolic rate for the $i$-th measurement of the $j$-th individual, $\alpha_j$ is the individual-specific intercept, $\beta_j$ is the individual-specific slope (thermal sensitivity), and $\sigma$ is the residual standard deviation.Because $T$ was centered at 25°C, $\alpha_j$ represents the log mass-specific metabolic rate of individual $j$ at that reference temperature.

In turn, the individual-specific parameters $\alpha_j$ and $\beta_j$ were modeled as being drawn from population-level distributions:

$$\alpha_j = \alpha + \alpha_S S_j + \alpha_Y Y_j + \alpha_M M_j + \eta_{\alpha,j}$$
$$\beta_j = \beta + \beta_S S_j + \beta_Y Y_j + \beta_M M_j + \eta_{\beta,j}$$

Where $S_j$, $Y_j$, and $M_j$ represent the site, year, and body mass of individual $j$, respectively, and $\eta_{\alpha,j}$ and $\eta_{\beta,j}$ are individual-level deviations from the population-level effects.

The individual-level deviations $\eta_{\alpha,j}$ and $\eta_{\beta,j}$ were modeled as being drawn from a multivariate normal distribution to account for the correlation between an individual's baseline metabolic rate and its thermal sensitivity:

$$\begin{bmatrix} \eta_{\alpha,j} \\ \eta_{\beta,j} \end{bmatrix} \sim \text{MVN} \left( \mathbf{0}, \mathbf{\Sigma} \right)$$

Where $\mathbf{0}$ is a vector of means and $\mathbf{\Sigma}$ is the covariance matrix, defined as:

$$\mathbf{\Sigma} = \begin{bmatrix} \tau_{\alpha}^2 & \rho\tau_{\alpha}\tau_{\beta} \\ \rho\tau_{\alpha}\tau_{\beta} & \tau_{\beta}^2 \end{bmatrix}$$

Where $\tau_{\alpha}$ and $\tau_{\beta}$ represent the among-individual standard deviations for the intercept and slope, respectively, and $\rho$ is the correlation coefficient between them.

#### Computational implementation

To ensure efficient sampling and avoid issues related to the geometry of the posterior distribution, we implemented the model using a non-centered parameterization. Instead of sampling the individual-level parameters $\alpha_j$ and $\beta_j$ directly from their respective group-level distributions, we treated them as transformed parameters. Specifically, we defined the vector of individual parameters $\mathbf{\Theta}_j = [\alpha_j, \beta_j]^T$ as:

$$\mathbf{\Theta}_j = \boldsymbol{\mu}_j + \text{diag}(\boldsymbol{\tau}) \mathbf{L}_{\Omega} \mathbf{z}_j$$

Where:

* $\boldsymbol{\mu}_j$ is the vector of population-level predicted means for the $j$-th individual.

* $\boldsymbol{\tau}$ is the vector of among-individual standard deviations ($\tau_{\alpha}, \tau_{\beta}$).

* $\mathbf{L}_{\Omega}$ is the Cholesky factor of the correlation matrix, representing the dependency between intercepts and slopes.

* $\mathbf{z}_j$ is a vector of standardized deviations sampled from a unit normal distribution, $\mathbf{z}_j \sim \mathcal{N}(0, 1)$.

#### Prior specification

We employed weakly informative priors for all model parameters to provide regularizing effects while ensuring that the posterior distributions remained primarily driven by the empirical data. We chose to center the priors for the global intercept ($\alpha$) and the effect sizes of site, year, and body mass ($\alpha_S, \alpha_Y, \alpha_M, \beta_S, \beta_Y, \beta_M$) at zero ($\mathcal{N}(0, 0.5)$). We adopted this approach as a conservative baseline assumption, reflecting a null expectation of no effect for environmental or physical covariates in the absence of strong evidence.

For the global thermal sensitivity parameter ($\beta$), we applied a $\mathcal{N}(0.1, 0.5)$ prior truncated at zero. We justified this truncation based on the fundamental biological law that ectotherm metabolic rate increases with body temperature; the standard deviation of 0.5 was selected to allow the model to explore a wide range of biologically plausible thermal sensitivities.

Regarding the variance components, we assigned $\mathcal{N}(0, 0.5)$ priors (constrained to positive values) to the residual standard deviation ($\sigma$) and the among-individual standard deviations ($\tau_{\alpha}, \tau_{\beta}$). To model the dependency between individual-level intercepts and slopes, we applied an $\text{LKJ}(2)$ prior to the Cholesky factor of the correlation matrix ($\mathbf{L}_{\Omega}$). We selected a shape parameter of 2 to provide a weakly informative prior that is intentionally skeptical of extreme correlations (near $\pm 1$), thereby favoring a more conservative estimate of the relationship between an individual’s baseline metabolism and its thermal sensitivity.

#### Inference and diagnostics

### Code implementation

#### `stan` code and model compilation

Below is the `stan` code for the resting metabolic rates model:

```{r print-model-code}
#| class-output: stan
#| code-fold: false
#| echo: false

cat(readLines("stan/resting-metabolic-rates.stan"), sep = "\n")

```

We pre-compute the `stan` model and save it as an RDS file to facilitate faster model fitting in subsequent analyses.

```{r stan-model-compilation}
#| eval: false
#| code-summary: "Compile `stan` model and save as RDS file"

## Note: This code chunk is not evaluated to avoid long rendering times.
## The compiled model is saved as an RDS file for later use.

rmr_model <- stan_model("stan/resting-metabolic-rates.stan")
saveRDS(rmr_model, "stan/resting-metabolic-rates.rds")
```

#### Data preparation

We prepare the data for `stan` model fitting. This involves filtering out rows with missing values, recoding site information, and arranging the data by year and individual ID. We then create a list containing the necessary variables for the `stan` model. In particular, `T` (body temperature) is centered around 25 degrees and scaled by 10 as we measured four temperatures (10, 20, 30, 40) and 25 represents the middle point of the range. We also scale `log_rmr_mass_specific` and `mass` to have mean 0 and standard deviation 1.

```{r stan-data-preparation}
#| code-fold: false
#| code-summary: "Prepare data for `stan` model fitting"

# filter out rows with missing log_rmr_mass_specific values
rmr_model_data <- rmr_data |>
  filter(!is.na(log_rmr_mass_specific)) |>
  mutate(
    site = paste(site, "E", sep = ""),
    site = ifelse(site == "TLE", "LE", site) # recode TLE to LE
  ) |>
  arrange(year, id)

# get list of unique site, year, individual combination
lizard_ids <- rmr_model_data |> select(id, site, year) |> distinct()

# pepare data as list
rmr_stan_data <- list(
  N = nrow(rmr_model_data),
  J = nrow(lizard_ids),
  I = as.numeric(as.factor(paste(rmr_model_data$id, rmr_model_data$year))),
  T = (rmr_model_data$temp - 25) / 10,
  R = as.numeric(scale(rmr_model_data$log_rmr_mass_specific)),
  M = as.numeric(scale(rmr_model_data$mass)),
  S = as.numeric(as.factor(lizard_ids$site)),
  Y = as.numeric(as.factor(lizard_ids$year))
)
```

#### Model fitting

```{r stan-model-fitting}
#| eval: false
#| code-fold: false
#| code-summary: "Fit `stan` model to data"

## Note: This code chunk is not evaluated to avoid long rendering times.

# load compiled stan model
rmr_model <- readRDS("stan/resting-metabolic-rates.rds")

# define function to provide initial values within prior constraints
inits <- function() {
  list(
    a = -5,
    b = 0.1,
    Sigma = 0.5,
    Tau = c(0.2, 0.2)
  )
}

# sample from model
rmr_model_fit <- sampling(
  rmr_model,
  data = rmr_stan_data,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = 4,
  init = inits,
  control = list(adapt_delta = 0.95, max_treedepth = 15)
)

# save fitted model
saveRDS(rmr_model_fit, "raw_data/rmr_model_fit.rds")
```

:::{.callout-note}
The object `rmr_model_fit` contains the fitted model results. It is saved locally in `raw_data/rmr_model_fit.rds` due to its large size. However, for transparency, the model summary and posterior distributions are shown below
:::

```{r show-model-fit}
#| code-summary: "Show model fit"
#| fig.align: "center"
#| fig.width: 4
#| fig.height: 4
#| out-width: "66%"

# load model fit
rmr_model_fit <- readRDS("raw_data/rmr_model_fit.rds")

print(
  rmr_model_fit,
  pars = c("a", "b", "a_S", "a_Y", "a_M", "b_S", "b_Y", "b_M")
)

mcmc_areas(
  rmr_model_fit,
  pars = c("a", "b", "a_S", "a_Y", "a_M", "b_S", "b_Y", "b_M"),
  prob = 0.95
) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Estimated value", y = "Parameter"
  )
```

In the plot above the parameter correspondance is as follows:

- `a` is $\alpha$: Overall intercept
- `b` is $\beta$: Overall slope
- `a_S` is $\alpha_S$: Site-level intercept
- `a_Y` is $\alpha_Y$: Year-level intercept
- `a_M` is $\alpha_M$: Mass-level intercept
- `b_S` is $\beta_S$: Site-level slope
- `b_Y` is $\beta_Y$: Year-level slope
- `b_M` is $\beta_M$: Mass-level slope

#### Model diagnostics

In addition to the effective sample sizes (`n_eff`) and R-hat values (`Rhat`) of each parameter we present above, below we also assess model quality via 3 common diagnostics: trace plots, posterior predictive checks, and the examination of residuals.

:::{.panel-tabset}

## HMC diagnostics

```{r run-hmc-diagnostics}
#| code-fold: false
check_hmc_diagnostics(rmr_model_fit)
```

## Trace plots

```{r trace-plots}
#| fig.align: "center"
#| fig.width: 6
#| fig.height: 4
#| code-fold: false
mcmc_trace(
  rmr_model_fit,
  pars = c("a", "b", "a_S", "a_Y", "a_M", "b_S", "b_Y", "b_M")
)
```

## Posterior predictive checks

```{r posterior-predictive-checks}
#| fig.align: "center"
#| fig.width: 6
#| fig.height: 4
#| out-width: "66%"
#| code-fold: false

# extract predicted values from model fit
R_pred <- extract(rmr_model_fit)$R_pred

# plot density overlay
ppc_dens_overlay(y = rmr_stan_data$R, yrep = R_pred[1:100, ]) +
labs(x = "Log mass-specific metabolic rate (scaled)", y = "Density")
```


## Examination of residuals

```{r residuals}
#| fig.align: "center"
#| fig.width: 6
#| fig.height: 4
#| out-width: "66%"

rmr_posteriors <- extract(rmr_model_fit)$R_pred
rmr_data_comparison <- tibble(
  body_temp = (rmr_stan_data$T * 10) + 25, # back-transform to original scale
  observed_rmr = rmr_stan_data$R,
  id = paste(rmr_model_data$id, rmr_model_data$year),
  predicted_rmr_mean = apply(rmr_posteriors, 2, mean),
  predicted_rmr_lower = apply(rmr_posteriors, 2, quantile, probs = 0.025),
  predicted_rmr_upper = apply(rmr_posteriors, 2, quantile, probs = 0.975)
)
ggplot(rmr_data_comparison, aes(x = observed_rmr, y = predicted_rmr_mean)) +
  geom_linerange(
    aes(ymin = predicted_rmr_lower, ymax = predicted_rmr_upper),
    alpha = 0.25
  ) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = 2) +
  labs(x = "Observed RMR", y = "Predicted RMR (mean ± 95% CI)")
```

:::

### Extracting individual-level paramters

```{r extracting-individual-level-parameters}
#| code-summary: "Extracting individual-level parameters from the model"
#| code-fold: false

# extract individual-level parameters from model
rmr_params <- rmr_model_fit |>
  spread_draws(a_I[J], b_I[J]) |>
  group_by(J) |>
  summarize(
    mean_intercept = mean(a_I),
    mean_slope = mean(b_I)
  )

# re-scale parameters and add individual IDs
rmr_mean <- mean(rmr_model_data$log_rmr_mass_specific, na.rm = TRUE)
rmr_sd <- sd(rmr_model_data$log_rmr_mass_specific, na.rm = TRUE)
rmr_params <- rmr_params |>
  mutate(
    slope = (mean_slope * rmr_sd) / 10,
    intercept_25 = (mean_intercept * rmr_sd) + rmr_mean,
    intercept = intercept_25 - slope * 25,
    site = lizard_ids$site[J],
    id = lizard_ids$id[J],
    year = lizard_ids$year[J]
  ) |>
  mutate(site = ifelse(startsWith(id, "T"), "TLE", site)) |>
  select(id, year, site, intercept, slope) |>
  unique()
```

```{r plot-individual-parameters}
#| code-summary: "Plotting individual-level parameters"
#| fig.align: "center"
#| fig.width: 7
#| fig.height: 4
#| out-width: "100%"

rmr_params |>
  pivot_longer(
    cols = c(intercept, slope),
    names_to = "parameter",
    values_to = "value"
  ) |>
  mutate(population = paste(site, year)) |>
  filter(population != "HE 2024") |>
  ggplot(aes(x = paste(site, year), y = value, color = population)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +
  stat_summary(
    fun.data = mean_sdl, fun.args = list(mult = 1)
  ) +
  facet_wrap(
    ~parameter, scales = "free_y",
    labeller = as_labeller(
      c(
        intercept = "Intercept~of~log(RMR) %~% T[b]",
        slope = "Slope~of~log(RMR) %~% T[b]"
      ),
      default = label_parsed
    )
  ) +
  labs(x = "Population", y = "Parameter value") +
  scale_color_manual(
    values = population_colors
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12)
  )
```

## Relationship between individual-level intercepts and slopes

```{r intercept-slope-relationship}
#| code-summary: "Plotting the relationship between individual-level intercepts and slopes"
#| fig.align: "center"
#| fig.width: 6
#| fig.height: 4
#| out-width: "75%"

rmr_params |>
  ggplot(aes(x = intercept, y = slope, color = paste(site, year))) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Individual-level intercept",
    y = "Individual-level slope",
    color = "Population"
  ) +
  scale_color_manual(
    values = population_colors
  ) +
  theme(
    legend.position = "right",
    strip.text = element_text(size = 12)
  )
```

Across all populations there appears to be an extremely strong correlation between intercept and slope. This allows us to simplify each individual's metabolic reaction norm by creating a composite value that captures both traits at the same time. We call this variable the "metabolic reaction norm" (MRN) and we define it as:

$$
MRN_i = \frac{\beta_i - \overline{\beta}}{\operatorname{SD}(\beta)} - \frac{\alpha_i - \overline{\alpha}}{\operatorname{SD}(\alpha)}
$$

```{r defining-and-plotting-mrn}
#| code.summary: "Defining and plotting the metabolic reaction norm (MRN)"
#| fig.align: "center"
#| fig.width: 6
#| fig.height: 4

rmr_params <- rmr_params |>
  mutate(
    scaled_intercept = (intercept - mean(intercept)) / sd(intercept),
    scaled_slope = (slope - mean(slope)) / sd(slope)
  ) |>
  mutate(mrn = (scaled_slope - scaled_intercept))

rmr_params |>
  pivot_longer(
    cols = c(scaled_intercept, scaled_slope),
    names_to = "parameter",
    values_to = "value"
  ) |>
  ggplot(aes(x = value, y = mrn, color = paste(site, year))) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Parameter value",
    y = "Metabolic Reaction Norm (MRN)",
    color = "Population"
  ) +
  scale_color_manual(
    values = population_colors
  ) +
  facet_wrap(
    ~parameter, scales = "free_x", axes = "all", strip.position = "bottom",
    labeller = as_labeller(c(
      scaled_intercept = "Scaled Intercept (α)",
      scaled_slope = "Scaled Slope (β)"
    ))
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12),
    strip.placement = "outside",
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 12),
    panel.border = element_rect(fill = NA)
  )
```

## Seasonal plasticity in RMR

```{r combining-mrn-with-respirometry-metadata}
#| fig-align: "center"
#| fig.width: 6
#| fig.height: 4
#| out.width: "66%"

# get trial metadata
trial_metadata <- respirometry |>
  select(site, year, id, doy) |>
  distinct() |>
  mutate(site = paste(site, "E", sep = "")) |>
  group_by(site, year, id) |>
  summarise(doy = mean(doy), .groups = "drop")

# merge trial metadata with rmr_params
rmr_params |>
  left_join(trial_metadata, by = c("site", "year", "id")) |>
  filter(year == 2024, site != "HE") |>
  mutate(
    date = as.Date(doy, origin = paste0(year, "-01-01"))
  ) |>
  ggplot(aes(x = date, y = mrn)) +
  geom_jitter(
    aes(color = paste(site, year)),
    width = 0.2, 
    height = 0, 
    size = 4,
    alpha = 0.5
  ) +
  geom_smooth(method = "lm", color = "black") +
  labs(
    x = "Date",
    y = "Metabolic Reaction Norm (MRN)",
    color = "Population"
  ) +
  scale_color_manual(
    values = population_colors
  )
```